Data Exploration

### 2024-02-08 analysis/documents/2-data-exploration.qmd
update,fixing issues

```{r}
#| label: image
image_data <- readRDS(here::here("data", "objects", "annotations_animals.rds")) |> 
  dplyr::select(id, source_file_with_dep,
                image_id, species:comment, camera_id,
                maker_notes_sequence:maker_notes_user_label, plot_id:ymd) |> 
  dplyr::rename(file_path = source_file_with_dep)
head(image_data)
```

@@

```{r}
#| label: camdep
cam_dep_data <- readRDS(here::here("data", "objects", "cams_sf_coords.rds")) |> 
  sf::st_drop_geometry() |> 
  dplyr::select(dplyr::ends_with("_id"), date_deployed, date_retrieved, longitude, latitude)
stopifnot(sum(table(cam_dep_data$camera_id) != 1) == 0)
```

@@

```{r, fig.height=20, fig.width=8}
#| label: camdep-plot
cam_dep_data |> 
  dplyr::mutate(date_deployed = as.Date(date_deployed),
                date_retrieved = as.Date(date_retrieved)) |> 
  dplyr::mutate(status = ifelse(date_deployed < date_retrieved, "valid", "invalid")) |> 
  dplyr::mutate(status = ifelse(is.na(status), "invalid", status)) |> 
  ggplot2::ggplot() +
  ggplot2::geom_segment(ggplot2::aes(x = date_deployed, y = camera_id, 
                                     xend = date_retrieved, yend = camera_id,
                                     color = status)) +
  ggplot2::geom_point(ggplot2::aes(x = date_deployed, y = camera_id, 
                                     color = status)) + 
    ggplot2::geom_point(ggplot2::aes(x = date_retrieved, y = camera_id, 
                                     color = status)) + 
  ggplot2::scale_color_discrete(type = c("#E69F00","#56B4E9")) +
  ggplot2::theme_bw()
```

>>>>

```{r}
#| label: image
image_data <- readRDS(here::here("data", "objects", "annotations_animals.rds")) |> 
  dplyr::select(id, source_file_with_dep,
                image_id, species:comment, camera_id,
                maker_notes_sequence:maker_notes_user_label, plot_id:ymd) |> 
  dplyr::mutate(site_id = sapply(image_id, \(x) strsplit(x, "_")[[1]][1]),
                date_time = lubridate::as_date(date_time)) |> 
  dplyr::rename(file_path = source_file_with_dep) |> 
  dplyr::filter(!is.na(date_time))

# Fixing error
# image_data[image_data$date_time==max(image_data$date_time, na.rm = T),] |> View()
image_data[image_data$ymd=="2024_Aug_19",]$date_time <- lubridate::as_date("2022-08-19")
image_data[image_data$ymd=="2024_Aug_19",]$ymd <- "2024_Aug_19"

head(image_data)
```

@@

```{r}
#| label: camdep
cam_dep_data <- readRDS(here::here("data", "objects", "cams_sf_coords.rds")) |> 
  sf::st_drop_geometry() |> 
  dplyr::select(dplyr::ends_with("_id"), date_deployed, date_retrieved, longitude, latitude) |> 
  dplyr::mutate(date_deployed = lubridate::as_date(date_deployed),
                date_retrieved = lubridate::as_date(date_retrieved)) 
stopifnot(sum(table(cam_dep_data$camera_id) != 1) == 0)
```

@@

By cameras

```{r, fig.height=20, fig.width=8}
#| label: camdep-plot

cam_dep_data_with_status <- cam_dep_data |> 
  dplyr::mutate(status = ifelse(date_deployed < date_retrieved, "valid", "invalid")) |> 
  dplyr::mutate(status = ifelse(is.na(status), "invalid", status))

cam_dep_data_with_status |> 
  ggplot2::ggplot() +
  ggplot2::geom_segment(ggplot2::aes(x = date_deployed, y = camera_id, 
                                     xend = date_retrieved, yend = camera_id,
                                     color = status)) +
  ggplot2::geom_point(ggplot2::aes(x = date_deployed, y = camera_id, 
                                   color = status)) + 
  ggplot2::geom_point(ggplot2::aes(x = date_retrieved, y = camera_id, 
                                   color = status)) + 
  ggplot2::scale_color_discrete(type = c("#E69F00","#56B4E9")) +
  ggplot2::theme_bw()
```

By sites

```{r, fig.height=40, fig.width=8}
cam_dep_data_with_status |> 
  ggplot2::ggplot() +
  ggplot2::geom_segment(ggplot2::aes(x = date_deployed, y = site_id,
                                     xend = date_retrieved, yend = site_id,
                                     color = status), alpha = 0.15, linewidth = 3) +
  ggplot2::geom_point(ggplot2::aes(x = date_deployed, y = site_id,
                                   color = status)) +
  ggplot2::geom_point(ggplot2::aes(x = date_retrieved, y = site_id,
                                   color = status)) +
  ggplot2::geom_text(ggplot2::aes(x = date_deployed-40, y = site_id,
                                  label = camera_id,
                                  color = status), size = 3) +
  ggplot2::scale_color_discrete(type = c("#E69F00","#56B4E9")) +
  ggplot2::theme_bw()
```

With image density

```{r, fig.height=40, fig.width=8}
cam_dep_data_with_status |> 
  ggplot2::ggplot(ggplot2::aes(x = date_deployed, y = reorder(site_id, 
                                                              date_deployed,
                                                              decreasing = TRUE))) +
  ggplot2::geom_segment(ggplot2::aes(xend = date_retrieved, yend = site_id,
                                     color = status), alpha = 0.15, linewidth = 3) +
  ggplot2::geom_point(ggplot2::aes(x = date_deployed, y = site_id,
                                   color = status)) +
  ggplot2::geom_point(ggplot2::aes(x = date_retrieved, y = site_id,
                                   color = status)) +

  ggplot2::geom_point(data = image_data,
                      ggplot2::aes(x = date_time, y = site_id), 
                      color = "firebrick", size = 0.5) +

  ggplot2::geom_text(ggplot2::aes(x = date_deployed-60, y = site_id,
                                  label = camera_id,
                                  color = status), size = 3) +
  ggplot2::scale_color_discrete(type = c("#E69F00","#56B4E9")) +
  ggplot2::theme_bw()
```

Sorted by deployment date

```{r, fig.height=40, fig.width=8}
cam_dep_data_with_status |> 
  ggplot2::ggplot(ggplot2::aes(x = date_deployed, y = reorder(camera_id, 
                                                              date_deployed,
                                                              decreasing = TRUE))) +
  ggplot2::geom_segment(ggplot2::aes(xend = date_retrieved, yend = camera_id,
                                     color = status), alpha = 0.15, linewidth = 3) +
  ggplot2::geom_point(ggplot2::aes(x = date_deployed, y = camera_id,
                                   color = status)) +
  ggplot2::geom_point(ggplot2::aes(x = date_retrieved, y = camera_id,
                                   color = status)) +

  ggplot2::geom_point(data = image_data,
                      ggplot2::aes(x = date_time, y = camera_id), 
                      color = "firebrick", size = 0.5) +

  ggplot2::geom_text(ggplot2::aes(x = date_deployed-60, y = camera_id,
                                  label = site_id,
                                  color = status), size = 3) +
  ggplot2::scale_color_discrete(type = c("#E69F00","#56B4E9")) +
  ggplot2::theme_bw()
```
